Pipeline Performance Figure Generation

### 2023-03-23 TestingPipeline/Scripts/comparing_methods_visualization.Rmd
Updating figures with current rules.

@@

vs2_vb_seqtype <- vs2_vb %>% 
  group_by(comb, comb_score, seqtype, max_score_group) %>%
  summarise(n = n())
```

@@

>>>>

comparing VS and VS2 viruses
```{r}
vs2_vs <- data.frame(vs2=viral_scores_two[,1],
                     vs=viral_scores_two[,2],
                     comb_score=viral_scores_two[,1]+viral_scores_two[,2],
                     seqtype=viruses$seqtype,
                     max_score_group=viruses$max_score_group)

vs2_vs$comb[vs2_vs$comb_score<1] <- "neither"
vs2_vs$comb[vs2_vs$vs2==0.5 & vs2_vs$vs==0.5] <- "both_low"
vs2_vs$comb[vs2_vs$vs2>=1 & vs2_vs$vs==0.5] <- "vs2_high_vs_low"
vs2_vs$comb[vs2_vs$vs2>=0.5 & vs2_vs$vs>=1] <- "vs2_low_vs_high"
vs2_vs$comb[vs2_vs$vs2>=1 & vs2_vs$vs==0] <- "vs2_only"
vs2_vs$comb[vs2_vs$vs2==0 & vs2_vs$vs>=1] <- "vs_only"
vs2_vs$comb[vs2_vs$vs2>=1 & vs2_vs$vs>=1] <- "both_high"

vs2_vs_seqtype <- vs2_vs %>% 
  group_by(comb, comb_score, seqtype) %>%
  summarise(n = n())
```

```{r}
table(vs2_vs$comb)
```

```{r}
ggplot(vs2_vs_seqtype, aes(y=n, x=comb,
                   fill=as.factor(comb_score))) +
  geom_bar(stat="identity", color="black") +
  theme_light() +
  coord_flip() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    legend.position = "bottom"
  ) +
  scale_fill_brewer(palette = "PuOr", ) +
  xlab("") +
  ylab("Number of Sequences") + 
  facet_grid(~seqtype, scales = "free")
```

@@

vs2_vb_seqtype <- vs2_vb %>% 
  group_by(comb, comb_score, seqtype) %>%
  summarise(n = n())
```

@@

```{r}
ggplot(vs2_vb_seqtype, aes(y=n, x=comb,
                   fill=as.factor(comb_score))) +
  geom_bar(stat="identity", color="black") +
  theme_light() +
  coord_flip() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    legend.position = "bottom"
  ) +
  scale_fill_brewer(palette = "PuOr", ) +
  xlab("") +
  ylab("Number of Sequences") + 
  facet_grid(~seqtype, scales = "free")
```

### 2023-03-23 TestingPipeline/Scripts/examining_false_positives.Rmd
New visualizations for SI.
Visualizing fragments analyses.
Updating figures with current rules.

@@

VIBRANT
    Quality == "Complete Circular": +1
    Quality == "High Quality Draft": +1
    Quality == "Medium Quality Draft": +1
    Quality == "Low Quality Draft" & provirus: +0.5

Virsorter2
    Viral >= 50: +0.5
    Viral >= 0.95: +0.5
    RNA >= 0.9: +1
    lavidaviridae >= 0.9: +1
    NCLDV >= 0.9: +1

Virsorter
    category ==  1,4: +1
    category == 2,5: +0.5

DeepVirFinder:
    Score >= 0.7: +0.5

Tuning - No Viral Signature:
    Kaiju_viral = "cellular organisms": -0.5
    If host_genes >50 and NOT provirus: -1 
    If viral_genes == 0 and host_genes >= 1: -1
    If 3*viral_genes <= host_genes and NOT provirus: -1
    If length > 50,000 and hallmark <=1: -1
    If length < 5000 and checkv completeness <= 75: -0.5

Tuning - Viral Signature:
    Kaiju_viral = "Viruses": +1
    If %unknown >= 75 and length < 50000: +0.5
    If %viral >= 50: +0.5
    Hallmark > 2: +1

@@

```{r getting_viral_set_1}
getting_viral_set_1 <- function(input_seqs,
                                include_vibrant=FALSE, 
                                include_virsorter2=FALSE,
                                include_deepvirfinder=FALSE,
                                include_tuning_viral=FALSE,
                                include_tuning_not_viral=FALSE,
                                include_virsorter=FALSE) {

@@

  if (include_vibrant) {
    keep_score[input_seqs$vibrant_quality=="complete circular"] <- keep_score[input_seqs$vibrant_quality=="complete circular"] + 1
    keep_score[input_seqs$vibrant_quality=="high quality draft"] <- keep_score[input_seqs$vibrant_quality=="high quality draft"] + 1
    keep_score[input_seqs$vibrant_quality=="medium quality draft"] <- keep_score[input_seqs$vibrant_quality=="medium quality draft"] + 1
    keep_score[input_seqs$vibrant_quality=="low quality draft" & input_seqs$provirus] <- keep_score[input_seqs$vibrant_quality=="low quality draft" & input_seqs$provirus] + 0.5
  }

  if (include_virsorter2) {
    keep_score[input_seqs$viral>=50] <- keep_score[input_seqs$viral>=50] + 0.5
    keep_score[input_seqs$viral>=95] <- keep_score[input_seqs$viral>=95] + 0.5
    keep_score[input_seqs$RNA>=0.9] <- keep_score[input_seqs$RNA>=0.9] + 1
    keep_score[input_seqs$lavidaviridae>=0.9] <- keep_score[input_seqs$lavidaviridae>=0.9] + 1
    keep_score[input_seqs$NCLDV>=0.9] <- keep_score[input_seqs$NCLDV>=0.9] + 1
  }

  if (include_virsorter) {
    keep_score[input_seqs$category==1] <- keep_score[input_seqs$category==1] + 1
    keep_score[input_seqs$category==2] <- keep_score[input_seqs$category==2] + 0.5
    keep_score[input_seqs$category==4] <- keep_score[input_seqs$category==4] + 1
    keep_score[input_seqs$category==5] <- keep_score[input_seqs$category==5] + 0.5
  }

  if (include_deepvirfinder) {
    keep_score[input_seqs$score>=0.7 & input_seqs$checkv_length<20000] <- keep_score[input_seqs$score>=0.7 & input_seqs$checkv_length<20000] + 0.5
    keep_score[input_seqs$score>=0.9 & input_seqs$checkv_length<20000] <- keep_score[input_seqs$score>=0.9 & input_seqs$checkv_length<20000] + 0.5
  }

  if (include_tuning_viral) {
    keep_score[input_seqs$Kaiju_Viral=="Viruses"] <- keep_score[input_seqs$Kaiju_Viral=="Viruses"] + 0.5
    keep_score[input_seqs$hallmark>2] <- keep_score[input_seqs$hallmark>2] + 1
    keep_score[input_seqs$percent_unknown>=75 & input_seqs$checkv_length<50000] <- keep_score[input_seqs$percent_unknown>=75 & input_seqs$checkv_length<50000] + 0.5
    keep_score[input_seqs$percent_viral>=50] <- keep_score[input_seqs$percent_viral>=50] + 0.5
  }

  if (include_tuning_not_viral) {
    keep_score[input_seqs$Kaiju_Viral=="cellular organisms"] <- keep_score[input_seqs$Kaiju_Viral=="cellular organisms"] - 0.5
    keep_score[input_seqs$checkv_host_genes>50 & !input_seqs$provirus] <- keep_score[input_seqs$checkv_host_genes>50 & !input_seqs$provirus] - 1
    keep_score[input_seqs$checkv_viral_genes==0 & input_seqs$checkv_host_genes>=1] <- keep_score[input_seqs$checkv_viral_genes==0 & input_seqs$checkv_host_genes>=1] - 1
    keep_score[((input_seqs$checkv_viral_genes*3) <= input_seqs$checkv_host_genes) & !input_seqs$provirus] <- keep_score[((input_seqs$checkv_viral_genes*3) <= input_seqs$checkv_host_genes) & !input_seqs$provirus] - 1 # consider accounting for RNA viruses
    keep_score[input_seqs$checkv_length>500000 & input_seqs$hallmark<=1] <- keep_score[input_seqs$checkv_length>500000 & input_seqs$hallmark<=1] - 1
    keep_score[input_seqs$checkv_completeness<=75 & input_seqs$checkv_length<=5000] <- keep_score[input_seqs$checkv_completeness<=75 & input_seqs$checkv_length<=5000] - 0.5 # helped with protist contamination
  }

@@

>>>>

```{r}
viruses$kaiju_match_ratio <- viruses$len/viruses$checkv_length
```

@@

@@

```{r getting_viral_set_1}
getting_viral_set_1 <- function(input_seqs,
                                include_vibrant=FALSE, 
                                include_virsorter2=FALSE,
                                include_deepvirfinder=FALSE,
                                include_tuning_viral=FALSE,
                                tv_1=T, tv_2=T, tv_3=T, tv_4=T,
                                include_tuning_not_viral=FALSE,
                                ntv_1=T, ntv_2=T, ntv_3=T, ntv_4=T, ntv_5=F,
                                include_virsorter=FALSE) {

@@

  if (include_vibrant) {
    keep_score[input_seqs$vibrant_quality=="high quality draft"] <- keep_score[input_seqs$vibrant_quality=="high quality draft"] + 1
    keep_score[input_seqs$vibrant_quality=="medium quality draft"] <- keep_score[input_seqs$vibrant_quality=="medium quality draft"] + 1
    keep_score[input_seqs$vibrant_quality=="low quality draft" & input_seqs$provirus] <- keep_score[input_seqs$vibrant_quality=="low quality draft" & input_seqs$provirus] + 0.5
  }

  if (include_virsorter2) {
    keep_score[input_seqs$max_score>=0.50] <- keep_score[input_seqs$max_score>=0.50] + 0.5
    keep_score[input_seqs$max_score>=0.95] <- keep_score[input_seqs$max_score>=0.95] + 0.5  
  }

  if (include_virsorter) {
    keep_score[input_seqs$category==1] <- keep_score[input_seqs$category==1] + 1
    keep_score[input_seqs$category==2] <- keep_score[input_seqs$category==2] + 0.5
    keep_score[input_seqs$category==3] <- keep_score[input_seqs$category==3] + 0.5
    keep_score[input_seqs$category==4] <- keep_score[input_seqs$category==4] + 1
    keep_score[input_seqs$category==5] <- keep_score[input_seqs$category==5] + 0.5 
    keep_score[input_seqs$category==6] <- keep_score[input_seqs$category==6] + 0.5 
  }

  if (include_deepvirfinder) {
     # add if DVF calls viral
    keep_score[(input_seqs$score>=0.9 & input_seqs$pvalue<=0.05) & input_seqs$checkv_length<20000] <- keep_score[(input_seqs$score>=0.9 & input_seqs$pvalue<=0.05) & input_seqs$checkv_length<20000] + 0.5
    keep_score[(input_seqs$score>=0.7 & input_seqs$pvalue<=0.05) & input_seqs$checkv_length<20000] <- keep_score[(input_seqs$score>=0.7 & input_seqs$pvalue<=0.05) & input_seqs$checkv_length<20000] + 0.5
  }

  if (include_tuning_viral) {
    keep_score <- tuning_viral(current_score=keep_score,
                               ins=input_seqs,
                         include_tv_1=tv_1,
                         include_tv_2=tv_2,
                         include_tv_3=tv_3,
                         include_tv_4=tv_4)
  }

  if (include_tuning_not_viral) {
    # tuning removal 
    keep_score <- tuning_not_viral(current_score=keep_score,
                               ins=input_seqs,
                         include_tv_1=ntv_1,
                         include_tv_2=ntv_2,
                         include_tv_3=ntv_3,
                         include_tv_4=ntv_4,
                         include_tv_5=ntv_5)

  }

@@

# Just looking at the protists getting wrong

```{r}
viruses$kaiju_match_ratio <- viruses$len/viruses$checkv_length
```

```{r}
ggplot(viruses, aes(y=kaiju_match_ratio, x=seqtype)) +
  geom_boxplot() +
  geom_point() +
  theme_bw() +
  facet_wrap(~seqtype, scales = "free") 
```

### 2023-03-23 TestingPipeline/Scripts/individual_rule_set_visualizations.Rmd
Updating figures with current rules.

pheatmap::pheatmap(table(vscore$seqtype, vscore$confusion_matrix)/c(rowSums(table(vscore$seqtype, vscore$confusion_matrix))),
                   color=c("black", viridis(1000)),
                   cluster_rows = F,
                   cluster_cols = F,
                   display_numbers = T,
                   number_color = "white", 
                   fontsize_number = 10,
                   number_format = "%.4f")
```

>>>>

pheatmap::pheatmap(table(vscore$seqtype, vscore$confusion_matrix)/c(rowSums(table(vscore$seqtype, vscore$confusion_matrix))),
                   color=c("black", viridis(1000)),
                   cluster_rows = F,
                   cluster_cols = F,
                   display_numbers = T,
                   number_color = "white", 
                   fontsize_number = 10,
                   number_format = "%.4f")
```
comparing each tool's score
```{r}
viruses$keep_score_ind_vs2 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
                                              include_vibrant = F,
                                              include_virsorter2 = T,
                                              include_tuning_viral = F,
                                              include_tuning_not_viral = F,
                                              include_virsorter = F)

viruses$keep_score_ind_tv <- getting_viral_set_1(viruses, include_deepvirfinder = F,
                                              include_vibrant = F,
                                              include_virsorter2 = T,
                                              include_tuning_viral = T,
                                              include_tuning_not_viral = F,
                                              include_virsorter = F)

viruses$keep_score_ind_dvf <- getting_viral_set_1(viruses, include_deepvirfinder = T,
                                              include_vibrant = F,
                                              include_virsorter2 = F,
                                              include_tuning_viral = F,
                                              include_tuning_not_viral = F,
                                              include_virsorter = F)

viruses$keep_score_ind_vb <- getting_viral_set_1(viruses, include_deepvirfinder = F,
                                              include_vibrant = T,
                                              include_virsorter2 = F,
                                              include_tuning_viral = F,
                                              include_tuning_not_viral = F,
                                              include_virsorter = F)

viruses$keep_score_ind_ntv <- getting_viral_set_1(viruses, include_deepvirfinder = F,
                                              include_vibrant = F,
                                              include_virsorter2 = F,
                                              include_tuning_viral = F,
                                              include_tuning_not_viral = T,
                                              include_virsorter = F)

viruses$keep_score_ind_vs <- getting_viral_set_1(viruses, include_deepvirfinder = F,
                                              include_vibrant = F,
                                              include_virsorter2 = F,
                                              include_tuning_viral = F,
                                              include_tuning_not_viral = F,
                                              include_virsorter = T)

viruses$true_virus <- "not"
viruses$true_virus[viruses$seqtype=="virus"] <- "virus"

viruses_long_scores <- viruses %>% 
  select(contains("keep_score_ind_"), seqtype) %>%
  pivot_longer(cols=contains("keep_score_"), 
               names_to="rule_combination",
               values_to="viral_score") %>% 
  mutate(viral_score=as.factor(round(viral_score))) %>%
  group_by(rule_combination, viral_score, seqtype) %>%
  summarise(n = n())
```

```{r}
pal_purple <- RColorBrewer::brewer.pal(4, "Purples")
pal_orange <- RColorBrewer::brewer.pal(3, "Oranges")

ggplot(viruses_long_scores, aes(y=n, x=rule_combination,
                   fill=viral_score)) +
  geom_bar(stat="identity", color="black") +
  theme_light() +
  coord_flip() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    legend.position = "bottom"
  ) +
  scale_fill_manual(name="",
                     values = alpha(c(rev(pal_orange), pal_purple), 0.8)) +
  xlab("") +
  ylab("Number of Sequences") + 
  facet_grid(~seqtype, scales = "free")
```

### 2023-03-23 TestingPipeline/Scripts/paper_figure_generation.Rmd
Updating figures with current rules.

@@

viruses$keep_score_vs2_ntv_2 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
                                              include_vibrant = F,
                                              include_virsorter2 = T,
                                              include_tuning_viral = F,
                                              ntv_1=T, ntv_2=T, ntv_3=F, ntv_4=F, ntv_5=F,
                                              include_tuning_not_viral = T,
                                              include_virsorter = F)

viruses$keep_score_vs2_ntv_3 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
                                              include_vibrant = F,
                                              include_virsorter2 = T,
                                              include_tuning_viral = F,
                                              ntv_1=T, ntv_2=T, ntv_3=T, ntv_4=F, ntv_5=F,
                                              include_tuning_not_viral = T,
                                              include_virsorter = F)

viruses$keep_score_vs2_ntv_4 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
                                              include_vibrant = F,
                                              include_virsorter2 = T,
                                              include_tuning_viral = F,
                                              ntv_1=T, ntv_2=T, ntv_3=T, ntv_4=T, ntv_5=F,
                                              include_tuning_not_viral = T,
                                              include_virsorter = F)

@@

>>>>

same idea but VS2 base
```{r}
viruses$keep_score_vs2_tv_all <- getting_viral_set_1(viruses, include_deepvirfinder = F,
                                              include_vibrant = F,
                                              include_virsorter2 = T,
                                              include_tuning_viral = T,
                                              include_tuning_not_viral = F,
                                              include_virsorter = F)

viruses$keep_score_vs2_tv_4_tv_2 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
                                              include_vibrant = F,
                                              include_virsorter2 = T,
                                              include_tuning_viral = T,
                                              tv_1=F, tv_2=T, tv_3=F, tv_4=T,
                                              include_tuning_not_viral = F,
                                              include_virsorter = F)

viruses$keep_score_vs2_tv_4_tv_2_tv_1 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
                                              include_vibrant = F,
                                              include_virsorter2 = T,
                                              include_tuning_viral = T,
                                              tv_1=T, tv_2=T, tv_3=F, tv_4=T,
                                              include_tuning_not_viral = F,
                                              include_virsorter = F)

viruses$keep_score_vs2_tv_1 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
                                              include_vibrant = F,
                                              include_virsorter2 = T,
                                              include_tuning_viral = T,
                                              tv_1=T, tv_2=F, tv_3=F, tv_4=F,
                                              include_tuning_not_viral = F,
                                              include_virsorter = F)

viruses$keep_score_vs2_tv_2 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
                                              include_vibrant = F,
                                              include_virsorter2 = T,
                                              include_tuning_viral = T,
                                              tv_1=F, tv_2=T, tv_3=F, tv_4=F,
                                              include_tuning_not_viral = F,
                                              include_virsorter = F)

viruses$keep_score_vs2_tv_3 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
                                              include_vibrant = F,
                                              include_virsorter2 = T,
                                              include_tuning_viral = T,
                                              tv_1=F, tv_2=F, tv_3=T, tv_4=F,
                                              include_tuning_not_viral = F,
                                              include_virsorter = F)

viruses$keep_score_vs2_tv_4 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
                                              include_vibrant = F,
                                              include_virsorter2 = T,
                                              include_tuning_viral = T,
                                              tv_1=F, tv_2=F, tv_3=F, tv_4=T,
                                              include_tuning_not_viral = F,
                                              include_virsorter = F)

viruses$keep_score_vs2_tv_0 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
                                              include_vibrant = F,
                                              include_virsorter2 = T,
                                              include_tuning_viral = F,
                                              include_tuning_not_viral = F,
                                              include_virsorter = F)

viruses_long_scores <- viruses %>% 
  select(contains("keep_score_vs2_tv_"), seqtype) %>%
  pivot_longer(cols=contains("keep_score_"), 
               names_to="rule_combination",
               values_to="viral_score") %>% 
  mutate(viral_score=as.factor(round(viral_score))) %>%
  group_by(rule_combination, viral_score, seqtype) %>%
  summarise(n = n())
```

```{r}
pal_purple <- RColorBrewer::brewer.pal(4, "Purples")
pal_orange <- RColorBrewer::brewer.pal(1, "Oranges")

ggplot(viruses_long_scores, aes(y=n, x=rule_combination,
                   fill=viral_score)) +
  geom_bar(stat="identity", color="black") +
  theme_light() +
  coord_flip() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    legend.position = "bottom"
  ) +
  scale_fill_manual(name="",
                     values = alpha(c(rev(pal_orange[1]), pal_purple), 0.8)) +
  xlab("") +
  ylab("Number of Sequences") + 
  facet_grid(~seqtype, scales = "free")
```

@@

viruses$keep_score_vs2_ntv_2 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
                                              include_vibrant = F,
                                              include_virsorter2 = T,
                                              include_tuning_viral = F,
                                              ntv_1=F, ntv_2=T, ntv_3=F, ntv_4=F, ntv_5=F,
                                              include_tuning_not_viral = T,
                                              include_virsorter = F)

viruses$keep_score_vs2_ntv_3 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
                                              include_vibrant = F,
                                              include_virsorter2 = T,
                                              include_tuning_viral = F,
                                              ntv_1=F, ntv_2=F, ntv_3=T, ntv_4=F, ntv_5=F,
                                              include_tuning_not_viral = T,
                                              include_virsorter = F)

viruses$keep_score_vs2_ntv_4 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
                                              include_vibrant = F,
                                              include_virsorter2 = T,
                                              include_tuning_viral = F,
                                              ntv_1=F, ntv_2=F, ntv_3=F, ntv_4=T, ntv_5=F,
                                              include_tuning_not_viral = T,
                                              include_virsorter = F)

viruses$keep_score_vs2_ntv_5 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
                                              include_vibrant = F,
                                              include_virsorter2 = T,
                                              include_tuning_viral = F,
                                              ntv_1=F, ntv_2=F, ntv_3=F, ntv_4=F, ntv_5=T,
                                              include_tuning_not_viral = T,
                                              include_virsorter = F)

@@

viruses$keep_score_vs2_ntv_0_tv <- getting_viral_set_1(viruses, include_deepvirfinder = F,
                                              include_vibrant = F,
                                              include_virsorter2 = T,
                                              include_tuning_viral = T,
                                              ntv_1=F, ntv_2=F, ntv_3=F, ntv_4=F, ntv_5=F,
                                              include_tuning_not_viral = F,
                                              include_virsorter = F)

viruses$keep_score_vs2_ntv_all_tv <- getting_viral_set_1(viruses, include_deepvirfinder = F,
                                              include_vibrant = F,
                                              include_virsorter2 = T,
                                              include_tuning_viral = T,
                                              include_tuning_not_viral = T,
                                              include_virsorter = F)

viruses$keep_score_vs2_ntv_all_tv_4_2_1 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
                                              include_vibrant = F,
                                              include_virsorter2 = T,
                                              include_tuning_viral = T,
                                              tv_1=T, tv_2=T, tv_3=F, tv_4=T,
                                              include_tuning_not_viral = T,
                                              include_virsorter = F)
